---
title: "FCE_MAP Data Submission_EDI"
author: "Mack White"
format: html
editor: source
---

# Generation of *MAP* Data Submission for FCE

## Setup
```{r}
#| warning: false

library(readr)
library(tidyverse)
library(leaflet)
library(DT)
library(janitor)
library(stringr)
```

```{r}
#| warning: false

prior_submission <- read_csv("../csv outputs/LT_TDCS_Rehage_001_2021.csv")
glimpse(prior_submission)
```

### Current Columns in Data Set
#### ID *`(chr)`*
#### Year *`(dbl)`*
#### Date *`(date)`*
#### SEASON *`(chr)`*
#### River *`(chr)`*
#### Creek *`(dbl)`*
#### SITETYPE *`(chr)`*
#### BOUT *`(dbl)`*
#### A ton of species-level codes that are not very straightforward *`(dbl)`*
#### CPUE *`(dbl)`*
#### Distance *`(dbl)`*
#### CPUE100m *`(dbl)`*
#### DEPTH_Creek *`(dbl)`*
#### TEMP *`(dbl)`*
#### DO *`(dbl)`*
#### Salinity *`(dbl)`*
#### SH1DSLD *`(dbl)`*
#### SH1AVE15D *`(dbl)`*
#### UPDISTANCE *`(dbl)`*
#### MEANDIST *`(dbl)`*

#### The current data looks janky, column names are not very informative, overall dataset is not very *F.A.I.R*
#### Going to fix that this year (November 2023)

## Read in up-to-date MAP data, set up columns, and select what we need
```{r}
raw_dat <- read_csv("../data/MAP_yrs1thru19.csv")
glimpse(raw_dat)

dat <- raw_dat |> 
  mutate(ID = as.character(paste(Year, SEASON, DRAINAGE, SITE, BOUT, sep = "")),
         DATE = as.Date(paste(s.yr, s.mo, s.day,sep = "-")),
         Hydrologic_Year = as.character(HYDROYEAR),
         River = as.character(DRAINAGE),
         common_name = (str_replace(common_name, fixed(" "), "_")),
         Creek = as.factor(SITE)) |> 
      select(ID, Year, DATE, SEASON, River, Creek, BOUT, common_name, CATCHNUMBER, Distance, TRIPDEPTH, DOMGL, SALINITY, TempC)

glimpse(dat)
```

## Clean up the column names
```{r}
dat_clean <- dat |> 
      clean_names(case = "upper_camel", abbreviations = c("ID", "DOMGL"))

dat_sum <- dat_clean |> 
      group_by(ID, CommonName) |> 
      mutate(CatchSum = sum(Catchnumber)) |> 
      ungroup() |> 
      group_by(ID) |> 
      mutate(Catch = sum(Catchnumber),
             CPUE = (Catch/Distance)*100) |> 
      distinct()

glimpse(dat_sum)
```

## Convert from long to wide format
```{r}

dat_wide <- dat_sum |> 
      pivot_wider(
            names_from = CommonName,
            values_from = CatchSum,
            values_fill = 0
      )

glimpse(dat_wide)

# dat_wide <- dat_wide |> 
#       mutate(Creek = as.character(Creek),
#              Distance = as.numeric(Distance))

dat_wide <- dat_wide |>
      mutate(Creek = as.character(Creek),
             Distance = as.numeric(Distance)) |>
      select(ID, Year, Date, Season, River, Creek, Bout, Distance, DOMGL, Salinity, TempC, Marsh_killifish, Golden_topminnow, Golden_shiner, Chain_pickerel, Florida_gar, Grass_pickerel, Sunfishes, Largemouth_bass, Lake_chubsucker, Seminole_killifish, Hogchoker, Tadpole_madtom, Warmouth, Bluegill, Redear, Spotted_sunfish, Dollar_sunfish, Bluespotted_sunfish, Snook, Tarpon, Striped_mullet, Gray_snapper, American_eel, Mayan_cichlid, Redfish, Coastal_shiner, Blue_tilapia, Mosquitofish, Tidewater_mojarra, Inland_silverside, Jack_Crevalle, Crested_goby, Peacock_eel, Sheepshead_minnow, Rainwater_killifish, Pike_killifish, Bluefin_killifish, Golden_shiner, Golden_topminnow, Brook_silverside, Jewel_Cichlid, Pinfish, Ladyfish, Peacock_bass, Atlantic_needlefish, Gulf_killifish, Tidewater_silverside, Spotted_seatrout, Diamond_killifish, Gafftopsail_catfish, Brown_Hoplo, Mozambique_tilapia, Timucu, Code_goby, Yellow_bullhead, Hardhead_catfish, Brown_bullhead, Bay_anchovy, Taillight_shiner, Barracuda, Oscar, Pirate_perch, Snakehead, Redfin_pickerel, Banded_Topminnow, Catch, CPUE)

glimpse(dat_wide)

```

```{r}
write_csv(dat_wide, "../csv outputs/LT_TDCS_Rehage_001_2023.csv")

check <- read_csv("../csv outputs/LT_TDCS_Rehage_001_2023.csv")
glimpse(check)
```

